---
title: "Final Project"
author: "Rashi Saxena, Ada Niu, and Golnaz Abrishami"
date: "11/26/2018"
output:
  html_document:
    fit_height: 4
    fit_width: 4
    number_sections: no
    theme: cosmo
    toc: yes
    toc_float:
      collapsed: no
  pdf_document:
    toc: yes
---

```{r,echo=FALSE}
library(psych)
library(pracma)
library(lattice)
library(ggplot2)
library(dplyr)
library(scatterplot3d)
#install.packages("car")
library(car)
```
# Data Cleaning 

1. Reading the file and creating model for Responde variable which is the time that people spend on social media per day.
```{r,echo=FALSE}
SM=read.csv("SM.csv")
str(SM)
```

2. Deleting Nulls
```{r,echo=FALSE}
dim(SM)
SM = SM[complete.cases(SM), ]
dim(SM)
```
3. Ploting Pairs Plot for comparation:

```{r,echo=FALSE}
pairs(HPD~age+NSM,data=SM, main = "Pair Plot",
      pch = 21,bg = c("red", "royalblue", "green3"))
```
Apperantly there is no linear relatioship, but we see potential outliers in Number of social media account that can impact our dataset.

```{r,echo=FALSE}
SM %>%
  ggplot(aes(factor(NSM),HPD))+
  geom_boxplot(outlier.colour="red",
                outlier.size=2) +
  xlab("Number of Social Media Account") +
  ylab("Hours Per Day") +
  ggtitle("Boxplot")
```

After Ploting a boxplot we see some anomaly in number of social media account besides the potenial outliers. 
So, we checked our dataset and figured out that this anomaly derives from incorrect data-a few people have written their phone numbers in the number of social media accounts column. We also notice a boxplot for 0 social media accounts, which could be puzzling. The reason for this is because when we sent our survey out, we had only accounted for Facebook, Instagram, Pinterest, and Snapchat as our social media platforms, but to some of our responders count WhatsApp, WeChat, and Reddit as social media accounts so they entered 0 for the the number of social media accounts we asked for, and entered the number of hours they spend on social media using other platforms. Let's clean our data. 

```{r,echo=FALSE}
for(i in c(nrow(SM):1)){
  if(SM$NSM[i]>10 || SM$NSM[i]==0 || SM$age[i]>90){
    SM=SM[-c(i),]
  }
}
dim(SM)
```

# Clean Data Visuals

```{r,echo=FALSE}
pairs(HPD~as.numeric(age)+NSM,data=SM, main = "Pair Plot -- Based on Gender",
      pch = 21,bg = c("red", "royalblue", "green3")[unclass(SM$Sex)])
```
Now we have a better observations. Let's build a model. 

Let's take a look at our cleaner version of the boxplot 
```{r,echo=FALSE}
SM %>%
  ggplot(aes(factor(NSM),HPD))+
  geom_boxplot(outlier.colour="red",
                outlier.size=2) +
  xlab("Number of Social Media Account") +
  ylab("Hours Per Day") +
  ggtitle("Boxplot")

```

# Model 1

```{r,echo=FALSE}
m1= lm(HPD~NSM+as.numeric(age),data=SM)
summary(m1)

```

We see a pretty low R squared, so let's see we have some influential outliers. So, we run a cook distance to figure out whtere outliers are influentioal or no. 

Maybe it is time for us to try a new model with the other variables available to us. 

# Model 2
```{r,echo=FALSE} 
m2 <- lm(HPD~as.numeric(age) + NSM + as.factor(gender) + as.factor(occupation)+ as.factor(marital_status) + as.factor(education), data =SM)
summary(m2)
```

Using this model, we see a slight improvement in our R sqaured, but also notice our pvalue increases, which shows that our current predictors do not have a relationship with number of hours on social media per day. Given the p values of the different variables, we notice that only age, number of social media per hours, and the gender(male specifically) shows a p value of less than .05. 

# Model 3

Here, let's use only age, gender, and number of social media accounts. 

```{r,echo=FALSE}

m3 <- lm(HPD~NSM+as.numeric(age)+as.factor(gender), data =SM)
summary(m3)
```
Using this model, we notice a slight jump in Rsquared, but also a very low p value. We can also notice that number of social media accounts, has a high p value, so it might be interesting to look at a model with just age and gender as the predictor variables. 

# Model 4

```{r,echo=FALSE}

m4 <- lm(HPD~ as.numeric(age)+as.factor(gender), data = SM) 

summary(m4)
```

Our R squared is still pretty low so let's check for any outliers, correlation between coefficients, and linear assumptions. 

# Analyzing 

```{r, echo=FALSE}
cook=cooks.distance(m4)
plot(cook)

```
According to plot we do not have any observation with coock distance greater than one. So, none of the outliers are influential and we can keep them.

Let's see if we have a way of finding some correlation betweek coefficients, meaning that two predictor variables are correlated. Let's check the variance inflation factors for this two predictors.

```{r, echo=FALSE}
vif(m4)
```

We notice that our variance influence factor values are close to 1, which means indicates that the standard deviation of the coefficients will remain stable with the inclusion in the regression equation of the other predictor variables. Therfore, we do not have any correlation between our predictor variables. 

Let's check for linear assumptions

```{r,echo=FALSE}
plot(m4)

```
We can eee inequality of the variance from the Residual plot. 

We can also see from the NOrmal Q-Q plot does not follow a normal distribution. 

Let's check visually if there is any linear relationship between the predictor variables and response variable. 
```{r, echo=FALSE}

HPD <- SM$HPD

age <- as.numeric(SM$age)

gender <- as.factor(SM$gender)

scatterplot3d(HPD,age,gender,  pch=16, highlight.3d=TRUE, main="3D Scatterplot")

````
From our 3D scatterplot, we can clearly see that there is no linear relationship between our predictor variables and response variable. 

# Conclusion 

Given our p-value, we can say that Age and Number of Social Media Accounts, are good predictors of Hours Spend on Social Media, but we can also say that given our significantly low adjusted R squared, they are not very useful. Let's look at the linear plots to see if our model meets the assumptions. 

Based on our R^sqaured, we can say that only 3% of the variation in Hours Spent on Social Media is explained by Age and Numbe of Social Media Accounts. 

We can say that there is a very weak linear relationship between our response and predictor variables. 








